Exploration

### 2018-02-26 exploration.Rmd
random subsampling option for testing

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
```{r sp_i}
set.seed(20)
mean_abund <- 5000
useLC <- TRUE
regPlots <- T
ipmPlots <- F
n.row <- ifelse(useLC, NA, 10)
n.col <- ifelse(useLC, NA, 20)
n.cell <- ifelse(useLC, n_distinct(lc.df$CellID), n.row*n.col)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-3, 1.3), # b1 + b2*z
                    s_x=c(-.5), # b1*x1 + ...
                    g_z=c(1, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(-1), #b1*x1 + ...
                    g_sig=1, # N(E, sig)
                    fl_z=c(-6.5, 1.9, -.2, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(0, -.01, 0), # b1*x1 + ...
                    seed_z=c(2.9, 0.05), # b1 + b2*z
                    seed_x=c(-1), # b1*x1 + ...
                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02))
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
if(useLC) {
  env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
                   temp2=temp^2,
                   prec=c(scale(lc.df$bio12_mean)),
                   prec2=prec^2,
                   pOpn=c(scale(lc.df$nlcd1_mean)),
                   pOth=c(scale(lc.df$nlcd2_mean)),
                   pDec=c(scale(lc.df$nlcd3_mean)),
                   pEvg=c(scale(lc.df$nlcd4_mean)),
                   pMxd=c(scale(lc.df$nlcd5_mean)),
                   x=lc.df$left,
                   y=lc.df$top)
} else {
  env.df <- tibble(temp=rnorm(n.cell),
                 temp2=temp^2,
                 prec=rnorm(n.cell),
                 prec2=prec^2,
                 x=rep(1:n.col, each=n.row),
                 y=rep(1:n.row, times=n.col)) 
}
```

@@

########
## loop through cells
########
for(i in 1:n.cell) {
  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
  lo <- min(d[[i]]$size)*0.5
  hi <- max(d[[i]]$size)*1.5
  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)

@@

  # Plots
  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.1, col="red")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.1, col="red")
      abline(a=0, b=1, col="blue", lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.1, col="red")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.1, col="red")
    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
    plot(y, stable.dist1, ylab="", type="l", lwd=2,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    lines(y, stable.dist2, lwd=2, lty=2)
    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xx <- seq(0, hi*.5, length.out=100)
      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
  }
  if(ipmPlots) {
    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
    abline(a=0, b=1, col="white", lty=2)
    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
  }
}
lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(lambdas ~ temp, data=lam.df, log="y")
plot(lambdas ~ temp2, data=lam.df, log="y")
plot(lambdas ~ prec, data=lam.df, log="y")
plot(lambdas ~ prec2, data=lam.df, log="y")
```

>>>>

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
```{r sp_i}
set.seed(20)
subsamp <- 100
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
                    s_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(-1, -.1, 1, -.1), #b1*x1 + ...
                    g_sig=1, # N(E, sig)
                    fl_z=c(-7.5, 1.2, -.2, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-1, -.01, 1, -.01), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z
                    seed_x=c(-1, -.1, 1, -.1), # b1*x1 + ...
                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02))
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.df <- tibble(temp=c(scale(lc.df$bio1_mean)),
                 temp2=temp^2,
                 prec=c(scale(lc.df$bio12_mean)),
                 prec2=prec^2,
                 pOpn=c(scale(lc.df$nlcd1_mean)),
                 pOth=c(scale(lc.df$nlcd2_mean)),
                 pDec=c(scale(lc.df$nlcd3_mean)),
                 pEvg=c(scale(lc.df$nlcd4_mean)),
                 pMxd=c(scale(lc.df$nlcd5_mean)),
                 x=lc.df$left,
                 y=lc.df$top)
env.df <- env.df[cells,]
```

@@

########
## loop through cells
########
for(i in 1:n.cell) {
  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
  lo <- min(d[[i]]$size)*0.8
  hi <- max(d[[i]]$size)*1.1
  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)

@@

  # Plots
  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.1, col="red")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.1, col="red")
      abline(a=0, b=1, col="blue", lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower:", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.1, col="red")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.1, col="red")
    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
    plot(y, stable.dist1, ylab="", type="l", lwd=2,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    lines(y, stable.dist2, lwd=2, lty=2)
    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xx <- seq(0, hi*.5, length.out=100)
      lines(xx, dnorm(xx, p$rcrt[1], p$rcrt[2]), col="red")
  }
  if(ipmPlots) {
    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
    abline(a=0, b=1, col="white", lty=2)
    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
  }
  if(i %% 100 == 0) cat("Finished cell", i, "\n")
}
lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(lambdas ~ temp, data=lam.df, log="y")
plot(lambdas ~ prec, data=lam.df, log="y")
plot(lambdas, lambdas2)
plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
```
